---
title: "Marketing_Analysis"
author: "Owais Khan"
date: "11/17/2021"
output: html_document
---


#### I can explore new functions or packages and implement them into analyses

install.packages("randomNames")
```{r echo=FALSE}
library(tidyverse)
library(lubridate)
library(randomNames)
library(generator)
```


#### I can import data from a variety of file types

```{r}
marketing_data <- read_csv("~/STA 518/R-for-data-science/data/marketing_campaign.csv")
```


```{r}
print(spec(marketing_data))
```


#### I can restructure information to be in a “tidy” format.

```{r}
#Eliminating unwanted variables
marketing_data <- marketing_data %>% select(ID:NumWebVisitsMonth)

#Calculating age of customers and the years since they're purchasing from given marketing data
marketing_data <- marketing_data %>%
  select(ID:NumWebVisitsMonth) %>%
  mutate(age = year(today()) - Year_Birth, customer_since = year(today()) - year(Dt_Customer))

#Removing all NA values from marketing dataset
cleaned_marketing_data <- na.omit(marketing_data)

#Generating customer data with unique ID's from marketing dataset
customer_data <- marketing_data %>%
  select(ID) %>%
  mutate(first_name = randomNames(nrow(marketing_data), which.names="first"),
         last_name = randomNames(nrow(marketing_data), which.names="last"),
         email = paste0(first_name, "_", last_name, "@gmail.com"),
         phone_number = r_phone_numbers(nrow(marketing_data)))
```


```{r}
#Removing all NA values from marketing dataset
cleaned_marketing_data <- na.omit(marketing_data)
```

#### I can combine information from multiple data sources.

```{r}
merged_data <- merge(x=cleaned_marketing_data,y=customer_data,by="ID")
```

#### I can restructure information to be in a “tidy” format.

```{r}
#Finding out contact information of top 10 customers with highest purchase of Meat products
meat_lovers <- meat_lovers <- merged_data %>%
  arrange(desc(MntMeatProducts)) %>%
  head(10) %>%
  select(ID, first_name, last_name, phone_number)

print(meat_lovers)

#Finding out contact information of top 10 customers with highest purchase of Sweet products
sweet_lovers <- meat_lovers <- merged_data %>%
  arrange(desc(MntSweetProducts)) %>%
  head(10) %>%
  select(ID, first_name, last_name, phone_number)

print(sweet_lovers)

#Finding contact information of top 100 lapsed customers (customer who hasn't returned to shop in a while)
lapsed_customers <- merged_data %>%
  arrange(desc(Recency)) %>%
  head(100) %>%
  mutate(no_purchase_since = Recency) %>%
  select(ID, first_name, last_name, phone_number, no_purchase_since)

print(lapsed_customers %>% head(10))

```

#### I can apply a function to groupings within a data source.

```{r}
#Get mean of salaries as per education
educational_income_mean <-group_by(merged_data, Education) %>% 
  summarise(mean_income = mean(Income)) %>% 
  arrange(desc(mean_income))

print(educational_income_mean)
```

#### I can create graphical displays of data that highlight key features.

```{r}
ggplot(data = merged_data) + 
  geom_point(mapping = aes(x = age, y = MntWines,  color = Marital_Status)) +
  labs(title = "Wines purchased by different Age Groups", x = "Age", y = "Number of wines purchased")

```


