---
title: "success_failure"
author: "Owais Khan"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(tidyverse)
library(reshape)
```

#### I can import data from a variety of sources.

```{r}
dataset <- read.csv(file = "~/STA 518/R-for-data-science/data/country_success_failure.txt", 
                    header = TRUE, sep = ",")
```

#### I can write comments that explain the “why” (Why did you choose this approach instead of an alternative? What else did you try that didn’t work?) of my code.


#### I can implement resampling methods to make conclusions about data.

Make the data wider by transforming column variable into row 
```{r}
#I had to transform the data into wider shape so that I can have access to each variable in the same row and I can perform the calculation simply instead of looping through each value and finding each variable.
wide_data <- pivot_wider(dataset, id_cols = c("ID", "NAME"), names_from = "variable", values_from = "estimate")
```

Rename variable ID to Id, NAME to Country, B01001_001 to Total, B01001_002 to Success & B01001_026 to Failure

```{r}
names(wide_data) <- c("Id" ,"Country", "Total", "Success", "Failure")
```

Calculate the success and failure rate from the respective variables.
```{r}
wide_data <- wide_data %>% mutate(Success_Rate = (Success/Total)*100, Failure_Rate = (Failure/Total)*100)
print(wide_data)
```

```{r}
#Mean success and failure rates throughout the countries
means <- wide_data %>%  summarise(mean_success = mean(Success_Rate), mean_failure = mean(Failure_Rate))
print(means)
```